---
title: "phylo_analysis"
author: "Ben Scott"
date: '2022-05-24'
output: html_document
---


```{r packages}
#### PGLS for-loop #### 
library(ape)
library(phytools)
library(Rcpp)
library(geiger)
library(caper)
library(MCMCglmm)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(MuMIn)

########################################################################

setwd("C:/Research/Masters_thesis/MT_tree_analysis")

# Read in Tree and data 
Card.tree <- read.nexus("FullMCC.tree.nexus")
data <- read.csv("Chap2data.csv", header = TRUE) 

rownames(data) <- data[,1,]
Brill <- log(data[c(13,14)])
data <- data[c(-13,-14)]
df <- cbind(data,Brill)

df<-df[Card.tree$tip.label, ]
comp.data<-comparative.data(Card.tree, df, names.col = "phylo", vcv=TRUE, warn.dropped=TRUE)

```

## Test pgls
```{r test_pgls}
# Comparing male and female plumage 
a <- pgls(data=comp.data,fPC1~mPC1,lambda="ML")
summary(a)
with(data, plot(fPC1~mPC1, xlab = "fPC1", ylab = "mPC1",
                main = "fPC1 vs \ mPC1"))
abline(a)

b <- pgls(data=comp.data,fPC2~mPC2,lambda="ML")
summary(b)
b$aicc

with(data, plot(fPC2~mPC2, xlab = "fPC2", ylab = "mPC2",
                main = "fPC2 vs \ mPC2"))
abline(b)
```
## Chi squared test

```{r chi_sqar}
# mydata %>%
#   select() %>%
#   colnames() %>%
#   combn(2) %>%
#   t() %>%
#   as_tibble() %>%
#   rowise %>%
#   mutate(chisq.test = list(
#     table(mydata[[nonforest]])
#   ))

chisq_results <- chisq.test(df$nonforest,df$Habitat) 
summary(chisq_results)
print(chisq_results)

chisq_results_2 <- chisq.test(df$nonforest,df$Strata) 
print(chisq_results_2)

chisq_results_3 <- chisq.test(df$nonforest,df$Forest.Dependency) 
print(chisq_results_3)

chisq_results_4 <- chisq.test(df$Forest.Dependency,df$Strata) 
print(chisq_results_4)

chisq_results_5 <- chisq.test(df$Forest.Dependency,df$Habitat) 
print(chisq_results_5)

chisq_results_6 <- chisq.test(df$Strata,df$Habitat) 
print(chisq_results_6)

chisq_results_7 <- chisq.test(df$Migration,df$Habitat) 
print(chisq_results_7)

chisq_results_8 <- chisq.test(df$Migration,df$Forest.Dependency) 
print(chisq_results_8)

```
```{r phylo_anova}
###### phylogenetic anova
# phylANOVA <- phylANOVA(Card.tree, comp.data$data$nonforest, comp.data$data$Habitat, nsim = 1000, posthoc=TRUE, p.adj="holm")
#x=aov.phylo(df$nonforest~df$Habitat, Card.tree, nsim=50, test="Wilks")
#print(attributes(x)$summary)

```
### nonforest and habitat are highly correlated, not independent. All traits except for 
### df$Forest.Dependency and df$Habitat are independent!! This means that my results for male color still hold water!



###### PGLS for-loop
# sET UP LISTS FOR LATER
```{r list set up}
TRAITS <- c("Openness",
            "Strata", "Migration","Forest Dependency","Habitat+Strata", "Openness+Migration",
            "Forest Dependency+Strata","Forest Dependency+Migration","Migration+Strata",
             "Migration+Strata+Openness","Migration+Strata+Forest Dependency","Habitat",
             "Habitat+Strata","Habitat+Migration","Migration+Strata+Habitat")
```
############################################
#https://github.com/enicurus/warbler.molt.migration/blob/master/mixed_models.R


```{r pgls_forloop}
# Just categorical variables 
mPC1mods<-list()
mPC1mods[[1]]<-pgls(data=comp.data,mPC1~nonforest,lambda="ML")
mPC1mods[[2]]<-pgls(data=comp.data,mPC1~Strata,lambda="ML")
mPC1mods[[3]]<-pgls(data=comp.data,mPC1~Migration,lambda="ML")
mPC1mods[[4]]<-pgls(data=comp.data,mPC1~Forest.Dependency,lambda="ML")
mPC1mods[[5]]<-pgls(data=comp.data,mPC1~nonforest+Strata,lambda="ML")
mPC1mods[[6]]<-pgls(data=comp.data,mPC1~nonforest+Migration,lambda="ML")
mPC1mods[[7]]<-pgls(data=comp.data,mPC1~Forest.Dependency+Strata,lambda="ML")
mPC1mods[[8]]<-pgls(data=comp.data,mPC1~Forest.Dependency+Migration,lambda="ML")
mPC1mods[[9]]<-pgls(data=comp.data,mPC1~Migration+Strata,lambda="ML")
mPC1mods[[10]]<-pgls(data=comp.data,mPC1~Migration+Strata+nonforest,lambda="ML")
mPC1mods[[11]]<-pgls(data=comp.data,mPC1~Migration+Strata+Forest.Dependency,lambda="ML")
mPC1mods[[12]]<-pgls(data=comp.data,mPC1~Habitat,lambda="ML")
mPC1mods[[13]]<-pgls(data=comp.data,mPC1~Habitat+Strata,lambda="ML")
mPC1mods[[14]]<-pgls(data=comp.data,mPC1~Habitat+Migration,lambda="ML")
mPC1mods[[15]]<-pgls(data=comp.data,mPC1~Migration+Strata+Habitat,lambda="ML")

# Collect summary statistics
mPC1modsAIC<-matrix(nrow=15,ncol=5)
for (i in 1:15){
  mPC1modsAIC[i,2]<-mPC1mods[[i]]$aicc
  mPC1modsAIC[i,3]<-summary(mPC1mods[[i]])$adj.r.squared
  mPC1modsAIC[i,4]<-pf(summary(mPC1mods[[i]])$fstatistic[1],summary(mPC1mods[[i]])$fstatistic[2],summary(mPC1mods[[i]])$fstatistic[3],lower.tail=F)
  mPC1modsAIC[i,5]<-mPC1mods[[i]]$param.CI$lambda$opt
  #mPC1modsAIC[i,6]<-summary(mPC1modsAIC[[i]])$coefficients[2]
  #mPC1modsAIC[i,7]<-summary(mPC1modsAIC[[i]])$coefficients[4]
  mPC1modsAIC[i,1]<-paste("Model",i,sep="_")
}

# transform to character
mPC1modsAIC<-data.frame(mPC1modsAIC)
mPC1modsAIC[,2]<-as.numeric(as.character(mPC1modsAIC[,2]))
mPC1modsAIC[,3]<-as.numeric(as.character(mPC1modsAIC[,3]))
mPC1modsAIC[,5]<-as.numeric(as.character(mPC1modsAIC[,5]))
mPC1modsAIC[,6]<-as.numeric(as.character(mPC1modsAIC[,6]))
mPC1modsAIC[,7]<-as.numeric(as.character(mPC1modsAIC[,7]))

# rename columns 
names(mPC1modsAIC)<-c("model","AICc","adj_r_sq","p_value","lambda", "Beta", "Std_Error")

# add plumage trait and habitat traits
mPC1modsAIC["Plumage_Trait"] <- c("mPC1")  ### change this out 
mPC1modsAIC["Habitat_Trait"] <- c(TRAITS)

mPC1modsAIC <- transform(mPC1modsAIC, `p_value` = as.numeric(`p_value`)) # convert p-value to numeric
mPC1modsAIC <- mPC1modsAIC %>% mutate(across(where(is.numeric), round, 4)) # round values to three decimal places

col_order <- c("Plumage_Trait", "Habitat_Trait", "model", "AICc", "adj_r_sq", "p_value", "lambda", "Beta", "Std_Error")
mPC1 <- mPC1modsAIC[, col_order]

mPC1<-mPC1[order(mPC1$AICc),]

write.csv(mPC1,"C:/Research/Masters_thesis/MT_tree_analysis/Results/PGLS/Chapter2/phylo_pca/mPC1_models.csv")

bonfor.mpc1 <- p.adjust(mPC1$`p-value`, method = p.adjust.methods, n = length(mPC1$`p-value`))


##############################
# Just categorical variables 
mPC2mods<-list()
mPC2mods[[1]]<-pgls(data=comp.data,mPC2~nonforest,lambda="ML")
mPC2mods[[2]]<-pgls(data=comp.data,mPC2~Strata,lambda="ML")
mPC2mods[[3]]<-pgls(data=comp.data,mPC2~Migration,lambda="ML")
mPC2mods[[4]]<-pgls(data=comp.data,mPC2~Forest.Dependency,lambda="ML")
mPC2mods[[5]]<-pgls(data=comp.data,mPC2~nonforest+Strata,lambda="ML")
mPC2mods[[6]]<-pgls(data=comp.data,mPC2~nonforest+Migration,lambda="ML")
mPC2mods[[7]]<-pgls(data=comp.data,mPC2~Forest.Dependency+Strata,lambda="ML")
mPC2mods[[8]]<-pgls(data=comp.data,mPC2~Forest.Dependency+Migration,lambda="ML")
mPC2mods[[9]]<-pgls(data=comp.data,mPC2~Migration+Strata,lambda="ML")
mPC2mods[[10]]<-pgls(data=comp.data,mPC2~Migration+Strata+nonforest,lambda="ML")
mPC2mods[[11]]<-pgls(data=comp.data,mPC2~Migration+Strata+Forest.Dependency,lambda="ML")
mPC2mods[[12]]<-pgls(data=comp.data,mPC2~Habitat,lambda="ML")
mPC2mods[[13]]<-pgls(data=comp.data,mPC2~Habitat+Strata,lambda="ML")
mPC2mods[[14]]<-pgls(data=comp.data,mPC2~Habitat+Migration,lambda="ML")
mPC2mods[[15]]<-pgls(data=comp.data,mPC2~Migration+Strata+Habitat,lambda="ML")

# Collect summary statistics
mPC2modsAIC<-matrix(nrow=15,ncol=5)
for (i in 1:15){
  mPC2modsAIC[i,2]<-mPC2mods[[i]]$aicc
  mPC2modsAIC[i,3]<-summary(mPC2mods[[i]])$adj.r.squared
  mPC2modsAIC[i,4]<-pf(summary(mPC2mods[[i]])$fstatistic[1],summary(mPC2mods[[i]])$fstatistic[2],summary(mPC2mods[[i]])$fstatistic[3],lower.tail=F)
  mPC2modsAIC[i,5]<-mPC2mods[[i]]$param.CI$lambda$opt
  mPC2modsAIC[i,1]<-paste("Model",i,sep="_")
}

# transform to character
mPC2modsAIC<-data.frame(mPC2modsAIC)
mPC2modsAIC[,2]<-as.numeric(as.character(mPC2modsAIC[,2]))
mPC2modsAIC[,3]<-as.numeric(as.character(mPC2modsAIC[,3]))
mPC2modsAIC[,5]<-as.numeric(as.character(mPC2modsAIC[,5]))

# rename columns 
names(mPC2modsAIC)<-c("model","AICc","adj_r_sq","p_value","lambda")

# add plumage trait and habitat traits
mPC2modsAIC["Plumage_Trait"] <- c("mPC2")  ### change this out 
mPC2modsAIC["Habitat_Trait"] <- c(TRAITS)

mPC2modsAIC <- transform(mPC2modsAIC, `p_value` = as.numeric(`p_value`)) # convert p-value to numeric
mPC2modsAIC <- mPC2modsAIC %>% mutate(across(where(is.numeric), round, 3)) # round values to three decimal places

col_order <- c("Plumage_Trait", "Habitat_Trait", "model", "AICc", "adj_r_sq", "p_value", "lambda")
mPC2 <- mPC2modsAIC[, col_order]

mPC2<-mPC2[order(mPC2$AICc),]

write.csv(mPC2,"C:/Research/Masters_thesis/MT_tree_analysis/Results/PGLS/Chapter2/phylo_pca/mPC2_models.csv")

################################
# Just categorical variables 
# Just categorical variables 
fPC1mods<-list()
fPC1mods[[1]]<-pgls(data=comp.data,fPC1~nonforest,lambda="ML")
fPC1mods[[2]]<-pgls(data=comp.data,fPC1~Strata,lambda="ML")
fPC1mods[[3]]<-pgls(data=comp.data,fPC1~Migration,lambda="ML")
fPC1mods[[4]]<-pgls(data=comp.data,fPC1~Forest.Dependency,lambda="ML")
fPC1mods[[5]]<-pgls(data=comp.data,fPC1~nonforest+Strata,lambda="ML")
fPC1mods[[6]]<-pgls(data=comp.data,fPC1~nonforest+Migration,lambda="ML")
fPC1mods[[7]]<-pgls(data=comp.data,fPC1~Forest.Dependency+Strata,lambda="ML")
fPC1mods[[8]]<-pgls(data=comp.data,fPC1~Forest.Dependency+Migration,lambda="ML")
fPC1mods[[9]]<-pgls(data=comp.data,fPC1~Migration+Strata,lambda="ML")
fPC1mods[[10]]<-pgls(data=comp.data,fPC1~Migration+Strata+nonforest,lambda="ML")
fPC1mods[[11]]<-pgls(data=comp.data,fPC1~Migration+Strata+Forest.Dependency,lambda="ML")
fPC1mods[[12]]<-pgls(data=comp.data,fPC1~Habitat,lambda="ML")
fPC1mods[[13]]<-pgls(data=comp.data,fPC1~Habitat+Strata,lambda="ML")
fPC1mods[[14]]<-pgls(data=comp.data,fPC1~Habitat+Migration,lambda="ML")
fPC1mods[[15]]<-pgls(data=comp.data,fPC1~Migration+Strata+Habitat,lambda="ML")

# Collect summary statistics
fPC1modsAIC<-matrix(nrow=15,ncol=5)
for (i in 1:15){
  fPC1modsAIC[i,2]<-fPC1mods[[i]]$aicc
  fPC1modsAIC[i,3]<-summary(fPC1mods[[i]])$adj.r.squared
  fPC1modsAIC[i,4]<-pf(summary(fPC1mods[[i]])$fstatistic[1],summary(fPC1mods[[i]])$fstatistic[2],summary(fPC1mods[[i]])$fstatistic[3],lower.tail=F)
  fPC1modsAIC[i,5]<-fPC1mods[[i]]$param.CI$lambda$opt
  fPC1modsAIC[i,1]<-paste("Model",i,sep="_")
}

# transform to character
fPC1modsAIC<-data.frame(fPC1modsAIC)
fPC1modsAIC[,2]<-as.numeric(as.character(fPC1modsAIC[,2]))
fPC1modsAIC[,3]<-as.numeric(as.character(fPC1modsAIC[,3]))
fPC1modsAIC[,5]<-as.numeric(as.character(fPC1modsAIC[,5]))

# rename columns 
names(fPC1modsAIC)<-c("model","AICc","adj_r_sq","p_value","lambda")

# add plumage trait and habitat traits
fPC1modsAIC["Plumage_Trait"] <- c("fPC1")  ### change this out 
fPC1modsAIC["Habitat_Trait"] <- c(TRAITS)

fPC1modsAIC <- transform(fPC1modsAIC, `p_value` = as.numeric(`p_value`)) # convert p-value to numeric
fPC1modsAIC <- fPC1modsAIC %>% mutate(across(where(is.numeric), round, 3)) # round values to three decimal places

col_order <- c("Plumage_Trait", "Habitat_Trait", "model", "AICc", "adj_r_sq", "p_value", "lambda")
fPC1 <- fPC1modsAIC[, col_order]

fPC1<-fPC1[order(fPC1$AICc),]

write.csv(fPC1,"C:/Research/Masters_thesis/MT_tree_analysis/Results/PGLS/Chapter2/phylo_pca/fPC1_models.csv")

##############################
# Just categorical variables 
##############################
# Just categorical variables 
fPC2mods<-list()
fPC2mods[[1]]<-pgls(data=comp.data,fPC2~nonforest,lambda="ML")
fPC2mods[[2]]<-pgls(data=comp.data,fPC2~Strata,lambda="ML")
fPC2mods[[3]]<-pgls(data=comp.data,fPC2~Migration,lambda="ML")
fPC2mods[[4]]<-pgls(data=comp.data,fPC2~Forest.Dependency,lambda="ML")
fPC2mods[[5]]<-pgls(data=comp.data,fPC2~nonforest+Strata,lambda="ML")
fPC2mods[[6]]<-pgls(data=comp.data,fPC2~nonforest+Migration,lambda="ML")
fPC2mods[[7]]<-pgls(data=comp.data,fPC2~Forest.Dependency+Strata,lambda="ML")
fPC2mods[[8]]<-pgls(data=comp.data,fPC2~Forest.Dependency+Migration,lambda="ML")
fPC2mods[[9]]<-pgls(data=comp.data,fPC2~Migration+Strata,lambda="ML")
fPC2mods[[10]]<-pgls(data=comp.data,fPC2~Migration+Strata+nonforest,lambda="ML")
fPC2mods[[11]]<-pgls(data=comp.data,fPC2~Migration+Strata+Forest.Dependency,lambda="ML")
fPC2mods[[12]]<-pgls(data=comp.data,fPC2~Habitat,lambda="ML")
fPC2mods[[13]]<-pgls(data=comp.data,fPC2~Habitat+Strata,lambda="ML")
fPC2mods[[14]]<-pgls(data=comp.data,fPC2~Habitat+Migration,lambda="ML")
fPC2mods[[15]]<-pgls(data=comp.data,fPC2~Migration+Strata+Habitat,lambda="ML")

# Collect summary statistics
fPC2modsAIC<-matrix(nrow=15,ncol=5)
for (i in 1:15){
  fPC2modsAIC[i,2]<-fPC2mods[[i]]$aicc
  fPC2modsAIC[i,3]<-summary(fPC2mods[[i]])$adj.r.squared
  fPC2modsAIC[i,4]<-pf(summary(fPC2mods[[i]])$fstatistic[1],summary(fPC2mods[[i]])$fstatistic[2],summary(fPC2mods[[i]])$fstatistic[3],lower.tail=F)
  fPC2modsAIC[i,5]<-fPC2mods[[i]]$param.CI$lambda$opt
  fPC2modsAIC[i,1]<-paste("Model",i,sep="_")
}

# transform to character
fPC2modsAIC<-data.frame(fPC2modsAIC)
fPC2modsAIC[,2]<-as.numeric(as.character(fPC2modsAIC[,2]))
fPC2modsAIC[,3]<-as.numeric(as.character(fPC2modsAIC[,3]))
fPC2modsAIC[,5]<-as.numeric(as.character(fPC2modsAIC[,5]))

# rename columns 
names(fPC2modsAIC)<-c("model","AICc","adj_r_sq","p_value","lambda")

# add plumage trait and habitat traits
fPC2modsAIC["Plumage_Trait"] <- c("fPC2")  ### change this out 
fPC2modsAIC["Habitat_Trait"] <- c(TRAITS)

fPC2modsAIC <- transform(fPC2modsAIC, `p_value` = as.numeric(`p_value`)) # convert p-value to numeric
fPC2modsAIC <- fPC2modsAIC %>% mutate(across(where(is.numeric), round, 3)) # round values to three decimal places

col_order <- c("Plumage_Trait", "Habitat_Trait", "model", "AICc", "adj_r_sq", "p_value", "lambda")
fPC2 <- fPC2modsAIC[, col_order]

fPC2<-fPC2[order(fPC2$AICc),]

write.csv(fPC2,"C:/Research/Masters_thesis/MT_tree_analysis/Results/PGLS/Chapter2/phylo_pca/fPC2_models.csv")
```

```{r brilliance pgls}
###########################################################################
## Brillance #########
# fAvgBrill
##############################
# Just categorical variables 
fAvgBrillmods<-list()
fAvgBrillmods[[1]]<-pgls(data=comp.data,fAvgBrill~nonforest,lambda="ML")
fAvgBrillmods[[2]]<-pgls(data=comp.data,fAvgBrill~Strata,lambda="ML")
fAvgBrillmods[[3]]<-pgls(data=comp.data,fAvgBrill~Migration,lambda="ML")
fAvgBrillmods[[4]]<-pgls(data=comp.data,fAvgBrill~Forest.Dependency,lambda="ML")
fAvgBrillmods[[5]]<-pgls(data=comp.data,fAvgBrill~nonforest+Strata,lambda="ML")
fAvgBrillmods[[6]]<-pgls(data=comp.data,fAvgBrill~nonforest+Migration,lambda="ML")
fAvgBrillmods[[7]]<-pgls(data=comp.data,fAvgBrill~Forest.Dependency+Strata,lambda="ML")
fAvgBrillmods[[8]]<-pgls(data=comp.data,fAvgBrill~Forest.Dependency+Migration,lambda="ML")
fAvgBrillmods[[9]]<-pgls(data=comp.data,fAvgBrill~Migration+Strata,lambda="ML")
fAvgBrillmods[[10]]<-pgls(data=comp.data,fAvgBrill~Migration+Strata+nonforest,lambda="ML")
fAvgBrillmods[[11]]<-pgls(data=comp.data,fAvgBrill~Migration+Strata+Forest.Dependency,lambda="ML")
fAvgBrillmods[[12]]<-pgls(data=comp.data,fAvgBrill~Habitat,lambda="ML")
fAvgBrillmods[[13]]<-pgls(data=comp.data,fAvgBrill~Habitat+Strata,lambda="ML")
fAvgBrillmods[[14]]<-pgls(data=comp.data,fAvgBrill~Habitat+Migration,lambda="ML")
fAvgBrillmods[[15]]<-pgls(data=comp.data,fAvgBrill~Migration+Strata+Habitat,lambda="ML")

# Collect summary statistics
fAvgBrillmodsAIC<-matrix(nrow=15,ncol=5)
for (i in 1:15){
  fAvgBrillmodsAIC[i,2]<-fAvgBrillmods[[i]]$aicc
  fAvgBrillmodsAIC[i,3]<-summary(fAvgBrillmods[[i]])$adj.r.squared
  fAvgBrillmodsAIC[i,4]<-pf(summary(fAvgBrillmods[[i]])$fstatistic[1],summary(fAvgBrillmods[[i]])$fstatistic[2],summary(fAvgBrillmods[[i]])$fstatistic[3],lower.tail=F)
  fAvgBrillmodsAIC[i,5]<-fAvgBrillmods[[i]]$param.CI$lambda$opt
  fAvgBrillmodsAIC[i,1]<-paste("Model",i,sep="_")
}

# transform to character
fAvgBrillmodsAIC<-data.frame(fAvgBrillmodsAIC)
fAvgBrillmodsAIC[,2]<-as.numeric(as.character(fAvgBrillmodsAIC[,2]))
fAvgBrillmodsAIC[,3]<-as.numeric(as.character(fAvgBrillmodsAIC[,3]))
fAvgBrillmodsAIC[,5]<-as.numeric(as.character(fAvgBrillmodsAIC[,5]))

# rename columns 
names(fAvgBrillmodsAIC)<-c("model","AICc","adj_r_sq","p_value","lambda")

# add plumage trait and habitat traits
fAvgBrillmodsAIC["Plumage_Trait"] <- c("fAvgBrill")  ### change this out 
fAvgBrillmodsAIC["Habitat_Trait"] <- c(TRAITS)

fAvgBrillmodsAIC <- transform(fAvgBrillmodsAIC, `p_value` = as.numeric(`p_value`)) # convert p-value to numeric
fAvgBrillmodsAIC <- fAvgBrillmodsAIC %>% mutate(across(where(is.numeric), round, 3)) # round values to three decimal places

col_order <- c("Plumage_Trait", "Habitat_Trait", "model", "AICc", "adj_r_sq", "p_value", "lambda")
fAvgBrill <- fAvgBrillmodsAIC[, col_order]

fAvgBrill<-fAvgBrill[order(fAvgBrill$AICc),]

write.csv(fAvgBrill,"C:/Research/Masters_thesis/MT_tree_analysis/Results/PGLS/Chapter2/phylo_pca/fAvgBrill_models.csv")
#############

# mAvgBrill
##############################
# Just categorical variables 
mAvgBrillmods<-list()
mAvgBrillmods[[1]]<-pgls(data=comp.data,mAvgBrill~nonforest,lambda="ML")
mAvgBrillmods[[2]]<-pgls(data=comp.data,mAvgBrill~Strata,lambda="ML")
mAvgBrillmods[[3]]<-pgls(data=comp.data,mAvgBrill~Migration,lambda="ML")
mAvgBrillmods[[4]]<-pgls(data=comp.data,mAvgBrill~Forest.Dependency,lambda="ML")
mAvgBrillmods[[5]]<-pgls(data=comp.data,mAvgBrill~nonforest+Strata,lambda="ML")
mAvgBrillmods[[6]]<-pgls(data=comp.data,mAvgBrill~nonforest+Migration,lambda="ML")
mAvgBrillmods[[7]]<-pgls(data=comp.data,mAvgBrill~Forest.Dependency+Strata,lambda="ML")
mAvgBrillmods[[8]]<-pgls(data=comp.data,mAvgBrill~Forest.Dependency+Migration,lambda="ML")
mAvgBrillmods[[9]]<-pgls(data=comp.data,mAvgBrill~Migration+Strata,lambda="ML")
mAvgBrillmods[[10]]<-pgls(data=comp.data,mAvgBrill~Migration+Strata+nonforest,lambda="ML")
mAvgBrillmods[[11]]<-pgls(data=comp.data,mAvgBrill~Migration+Strata+Forest.Dependency,lambda="ML")
mAvgBrillmods[[12]]<-pgls(data=comp.data,mAvgBrill~Habitat,lambda="ML", bounds = list(lambda=c(0.001,1), kappa=c(1e-6,3), delta=c(1e-6,3)))
mAvgBrillmods[[13]]<-pgls(data=comp.data,mAvgBrill~Habitat+Strata,lambda="ML")
mAvgBrillmods[[14]]<-pgls(data=comp.data,mAvgBrill~Habitat+Migration,lambda="ML")
mAvgBrillmods[[15]]<-pgls(data=comp.data,mAvgBrill~Migration+Strata+Habitat,lambda="ML")

# Collect summary statistics
mAvgBrillmodsAIC<-matrix(nrow=15,ncol=5)
for (i in 1:15){
  mAvgBrillmodsAIC[i,2]<-mAvgBrillmods[[i]]$aicc
  mAvgBrillmodsAIC[i,3]<-summary(mAvgBrillmods[[i]])$adj.r.squared
  mAvgBrillmodsAIC[i,4]<-pf(summary(mAvgBrillmods[[i]])$fstatistic[1],summary(mAvgBrillmods[[i]])$fstatistic[2],summary(mAvgBrillmods[[i]])$fstatistic[3],lower.tail=F)
  mAvgBrillmodsAIC[i,5]<-mAvgBrillmods[[i]]$param.CI$lambda$opt
  mAvgBrillmodsAIC[i,1]<-paste("Model",i,sep="_")
}

# transform to character
mAvgBrillmodsAIC<-data.frame(mAvgBrillmodsAIC)
mAvgBrillmodsAIC[,2]<-as.numeric(as.character(mAvgBrillmodsAIC[,2]))
mAvgBrillmodsAIC[,3]<-as.numeric(as.character(mAvgBrillmodsAIC[,3]))
mAvgBrillmodsAIC[,5]<-as.numeric(as.character(mAvgBrillmodsAIC[,5]))

# rename columns 
names(mAvgBrillmodsAIC)<-c("model","AICc","adj_r_sq","p_value","lambda")

# add plumage trait and habitat traits
mAvgBrillmodsAIC["Plumage_Trait"] <- c("mAvgBrill")  ### change this out 
mAvgBrillmodsAIC["Habitat_Trait"] <- c(TRAITS)

mAvgBrillmodsAIC <- transform(mAvgBrillmodsAIC, `p_value` = as.numeric(`p_value`)) # convert p-value to numeric
mAvgBrillmodsAIC <- mAvgBrillmodsAIC %>% mutate(across(where(is.numeric), round, 3)) # round values to three decimal places

col_order <- c("Plumage_Trait", "Habitat_Trait", "model", "AICc", "adj_r_sq", "p_value", "lambda")
mAvgBrill <- mAvgBrillmodsAIC[, col_order]

mAvgBrill<-mAvgBrill[order(mAvgBrill$AICc),]

write.csv(mAvgBrill,"C:/Research/Masters_thesis/MT_tree_analysis/Results/PGLS/Chapter2/phylo_pca/mAvgBrill_models.csv")
```
############### Taking the best models ###################

```{r best_model}

best_mpc1 <- mPC1[which.min(mPC1$AICc),]
best_mpc2 <- mPC2[which.min(mPC2$AICc),]
best_fpc1 <- fPC1[which.min(fPC1$AICc),]
best_fpc2 <- fPC2[which.min(fPC2$AICc),]
best_mbri <- mAvgBrill[which.min(mAvgBrill$AICc),]
best_fbri <- fAvgBrill[which.min(fAvgBrill$AICc),]

best_all_models <- rbind(best_mpc1,best_mpc2,best_fpc1,best_fpc2,best_mbri,best_fbri)
write_csv(best_all_models, file = "C:/Research/Masters_thesis/MT_tree_analysis/Results/PGLS/Chapter2/phylo_pca/best_models.csv")
```

### Run Identical code for non-phylogenetic PCA 

```{r nonphylo preprocessing}
df <- read.csv("C:/Research/Masters_thesis/MT_tree_analysis/Chap2data.nonphylo.csv", header = TRUE)
row.names(df) = df$Species

#get rid of node labels which prevent comparative dataframe object 
Card.tree$node.label<-NULL

comp.data<-comparative.data(Card.tree, df, names.col = "phylo", vcv=TRUE,vcv.dim = 3, warn.dropped=TRUE)
full <- comp.data$data

##############################################3
# sET UP LISTS FOR LATER
TRAITS <- c("Openness",
            "Strata", "Migration","Forest Dependency","Habitat+Strata", "Openness+Migration",
            "Forest Dependency+Strata","Forest Dependency+Migration","Migration+Strata",
             "Migration+Strata+Openness","Migration+Strata+Forest Dependency","Habitat",
             "Habitat+Strata","Habitat+Migration","Migration+Strata+Habitat")


```


```{r pgls_forloop}
# Just categorical variables 
mPC1mods<-list()
mPC1mods[[1]]<-pgls(data=comp.data,mPC1~nonforest,lambda="ML")
mPC1mods[[2]]<-pgls(data=comp.data,mPC1~Strata,lambda="ML")
mPC1mods[[3]]<-pgls(data=comp.data,mPC1~Migration,lambda="ML")
mPC1mods[[4]]<-pgls(data=comp.data,mPC1~Forest.Dependency,lambda="ML")
mPC1mods[[5]]<-pgls(data=comp.data,mPC1~nonforest+Strata,lambda="ML")
mPC1mods[[6]]<-pgls(data=comp.data,mPC1~nonforest+Migration,lambda="ML")
mPC1mods[[7]]<-pgls(data=comp.data,mPC1~Forest.Dependency+Strata,lambda="ML")
mPC1mods[[8]]<-pgls(data=comp.data,mPC1~Forest.Dependency+Migration,lambda="ML")
mPC1mods[[9]]<-pgls(data=comp.data,mPC1~Migration+Strata,lambda="ML")
mPC1mods[[10]]<-pgls(data=comp.data,mPC1~Migration+Strata+nonforest,lambda="ML")
mPC1mods[[11]]<-pgls(data=comp.data,mPC1~Migration+Strata+Forest.Dependency,lambda="ML")
mPC1mods[[12]]<-pgls(data=comp.data,mPC1~Habitat,lambda="ML")
mPC1mods[[13]]<-pgls(data=comp.data,mPC1~Habitat+Strata,lambda="ML")
mPC1mods[[14]]<-pgls(data=comp.data,mPC1~Habitat+Migration,lambda="ML")
mPC1mods[[15]]<-pgls(data=comp.data,mPC1~Migration+Strata+Habitat,lambda="ML")

# Collect summary statistics
mPC1modsAIC<-matrix(nrow=15,ncol=5)
for (i in 1:15){
  mPC1modsAIC[i,2]<-mPC1mods[[i]]$aicc
  mPC1modsAIC[i,3]<-summary(mPC1mods[[i]])$adj.r.squared
  mPC1modsAIC[i,4]<-pf(summary(mPC1mods[[i]])$fstatistic[1],summary(mPC1mods[[i]])$fstatistic[2],summary(mPC1mods[[i]])$fstatistic[3],lower.tail=F)
  mPC1modsAIC[i,5]<-mPC1mods[[i]]$param.CI$lambda$opt
  mPC1modsAIC[i,1]<-paste("Model",i,sep="_")
}

# transform to character
mPC1modsAIC<-data.frame(mPC1modsAIC)
mPC1modsAIC[,2]<-as.numeric(as.character(mPC1modsAIC[,2]))
mPC1modsAIC[,3]<-as.numeric(as.character(mPC1modsAIC[,3]))
mPC1modsAIC[,5]<-as.numeric(as.character(mPC1modsAIC[,5]))

# rename columns 
names(mPC1modsAIC)<-c("model","AICc","adj_r_sq","p_value","lambda")

# add plumage trait and habitat traits
mPC1modsAIC["Plumage_Trait"] <- c("mPC1")  ### change this out 
mPC1modsAIC["Habitat_Trait"] <- c(TRAITS)

mPC1modsAIC <- transform(mPC1modsAIC, `p_value` = as.numeric(`p_value`)) # convert p-value to numeric
mPC1modsAIC <- mPC1modsAIC %>% mutate(across(where(is.numeric), round, 4)) # round values to three decimal places

col_order <- c("Plumage_Trait", "Habitat_Trait", "model", "AICc", "adj_r_sq", "p_value", "lambda")
mPC1 <- mPC1modsAIC[, col_order]

mPC1<-mPC1[order(mPC1$AICc),]

write.csv(mPC1,"C:/Research/Masters_thesis/MT_tree_analysis/Results/PGLS/Chapter2/phylo_pca/mPC1_models.csv")

bonfor.mpc1 <- p.adjust(mPC1$`p-value`, method = p.adjust.methods, n = length(mPC1$`p-value`))


##############################
# Just categorical variables 
mPC2mods<-list()
mPC2mods[[1]]<-pgls(data=comp.data,mPC2~nonforest,lambda="ML")
mPC2mods[[2]]<-pgls(data=comp.data,mPC2~Strata,lambda="ML")
mPC2mods[[3]]<-pgls(data=comp.data,mPC2~Migration,lambda="ML")
mPC2mods[[4]]<-pgls(data=comp.data,mPC2~Forest.Dependency,lambda="ML")
mPC2mods[[5]]<-pgls(data=comp.data,mPC2~nonforest+Strata,lambda="ML")
mPC2mods[[6]]<-pgls(data=comp.data,mPC2~nonforest+Migration,lambda="ML")
mPC2mods[[7]]<-pgls(data=comp.data,mPC2~Forest.Dependency+Strata,lambda="ML")
mPC2mods[[8]]<-pgls(data=comp.data,mPC2~Forest.Dependency+Migration,lambda="ML")
mPC2mods[[9]]<-pgls(data=comp.data,mPC2~Migration+Strata,lambda="ML")
mPC2mods[[10]]<-pgls(data=comp.data,mPC2~Migration+Strata+nonforest,lambda="ML")
mPC2mods[[11]]<-pgls(data=comp.data,mPC2~Migration+Strata+Forest.Dependency,lambda="ML")
mPC2mods[[12]]<-pgls(data=comp.data,mPC2~Habitat,lambda="ML")
mPC2mods[[13]]<-pgls(data=comp.data,mPC2~Habitat+Strata,lambda="ML")
mPC2mods[[14]]<-pgls(data=comp.data,mPC2~Habitat+Migration,lambda="ML")
mPC2mods[[15]]<-pgls(data=comp.data,mPC2~Migration+Strata+Habitat,lambda="ML")

# Collect summary statistics
mPC2modsAIC<-matrix(nrow=15,ncol=5)
for (i in 1:15){
  mPC2modsAIC[i,2]<-mPC2mods[[i]]$aicc
  mPC2modsAIC[i,3]<-summary(mPC2mods[[i]])$adj.r.squared
  mPC2modsAIC[i,4]<-pf(summary(mPC2mods[[i]])$fstatistic[1],summary(mPC2mods[[i]])$fstatistic[2],summary(mPC2mods[[i]])$fstatistic[3],lower.tail=F)
  mPC2modsAIC[i,5]<-mPC2mods[[i]]$param.CI$lambda$opt
  mPC2modsAIC[i,1]<-paste("Model",i,sep="_")
}

# transform to character
mPC2modsAIC<-data.frame(mPC2modsAIC)
mPC2modsAIC[,2]<-as.numeric(as.character(mPC2modsAIC[,2]))
mPC2modsAIC[,3]<-as.numeric(as.character(mPC2modsAIC[,3]))
mPC2modsAIC[,5]<-as.numeric(as.character(mPC2modsAIC[,5]))

# rename columns 
names(mPC2modsAIC)<-c("model","AICc","adj_r_sq","p_value","lambda")

# add plumage trait and habitat traits
mPC2modsAIC["Plumage_Trait"] <- c("mPC2")  ### change this out 
mPC2modsAIC["Habitat_Trait"] <- c(TRAITS)

mPC2modsAIC <- transform(mPC2modsAIC, `p_value` = as.numeric(`p_value`)) # convert p-value to numeric
mPC2modsAIC <- mPC2modsAIC %>% mutate(across(where(is.numeric), round, 4)) # round values to three decimal places

col_order <- c("Plumage_Trait", "Habitat_Trait", "model", "AICc", "adj_r_sq", "p_value", "lambda")
mPC2 <- mPC2modsAIC[, col_order]

mPC2<-mPC2[order(mPC2$AICc),]

write.csv(mPC2,"C:/Research/Masters_thesis/MT_tree_analysis/Results/PGLS/Chapter2/phylo_pca/mPC2_models.csv")

################################
# Just categorical variables 
# Just categorical variables 
fPC1mods<-list()
fPC1mods[[1]]<-pgls(data=comp.data,fPC1~nonforest,lambda="ML")
fPC1mods[[2]]<-pgls(data=comp.data,fPC1~Strata,lambda="ML")
fPC1mods[[3]]<-pgls(data=comp.data,fPC1~Migration,lambda="ML")
fPC1mods[[4]]<-pgls(data=comp.data,fPC1~Forest.Dependency,lambda="ML")
fPC1mods[[5]]<-pgls(data=comp.data,fPC1~nonforest+Strata,lambda="ML")
fPC1mods[[6]]<-pgls(data=comp.data,fPC1~nonforest+Migration,lambda="ML")
fPC1mods[[7]]<-pgls(data=comp.data,fPC1~Forest.Dependency+Strata,lambda="ML")
fPC1mods[[8]]<-pgls(data=comp.data,fPC1~Forest.Dependency+Migration,lambda="ML")
fPC1mods[[9]]<-pgls(data=comp.data,fPC1~Migration+Strata,lambda="ML")
fPC1mods[[10]]<-pgls(data=comp.data,fPC1~Migration+Strata+nonforest,lambda="ML")
fPC1mods[[11]]<-pgls(data=comp.data,fPC1~Migration+Strata+Forest.Dependency,lambda="ML")
fPC1mods[[12]]<-pgls(data=comp.data,fPC1~Habitat,lambda="ML")
fPC1mods[[13]]<-pgls(data=comp.data,fPC1~Habitat+Strata,lambda="ML")
fPC1mods[[14]]<-pgls(data=comp.data,fPC1~Habitat+Migration,lambda="ML")
fPC1mods[[15]]<-pgls(data=comp.data,fPC1~Migration+Strata+Habitat,lambda="ML")

# Collect summary statistics
fPC1modsAIC<-matrix(nrow=15,ncol=5)
for (i in 1:15){
  fPC1modsAIC[i,2]<-fPC1mods[[i]]$aicc
  fPC1modsAIC[i,3]<-summary(fPC1mods[[i]])$adj.r.squared
  fPC1modsAIC[i,4]<-pf(summary(fPC1mods[[i]])$fstatistic[1],summary(fPC1mods[[i]])$fstatistic[2],summary(fPC1mods[[i]])$fstatistic[3],lower.tail=F)
  fPC1modsAIC[i,5]<-fPC1mods[[i]]$param.CI$lambda$opt
  fPC1modsAIC[i,1]<-paste("Model",i,sep="_")
}

# transform to character
fPC1modsAIC<-data.frame(fPC1modsAIC)
fPC1modsAIC[,2]<-as.numeric(as.character(fPC1modsAIC[,2]))
fPC1modsAIC[,3]<-as.numeric(as.character(fPC1modsAIC[,3]))
fPC1modsAIC[,5]<-as.numeric(as.character(fPC1modsAIC[,5]))

# rename columns 
names(fPC1modsAIC)<-c("model","AICc","adj_r_sq","p_value","lambda")

# add plumage trait and habitat traits
fPC1modsAIC["Plumage_Trait"] <- c("fPC1")  ### change this out 
fPC1modsAIC["Habitat_Trait"] <- c(TRAITS)

fPC1modsAIC <- transform(fPC1modsAIC, `p_value` = as.numeric(`p_value`)) # convert p-value to numeric
fPC1modsAIC <- fPC1modsAIC %>% mutate(across(where(is.numeric), round, 3)) # round values to three decimal places

col_order <- c("Plumage_Trait", "Habitat_Trait", "model", "AICc", "adj_r_sq", "p_value", "lambda")
fPC1 <- fPC1modsAIC[, col_order]

fPC1<-fPC1[order(fPC1$AICc),]

write.csv(fPC1,"C:/Research/Masters_thesis/MT_tree_analysis/Results/PGLS/Chapter2/phylo_pca/fPC1_models.csv")

##############################
# Just categorical variables 
##############################
# Just categorical variables 
fPC2mods<-list()
fPC2mods[[1]]<-pgls(data=comp.data,fPC2~nonforest,lambda="ML")
fPC2mods[[2]]<-pgls(data=comp.data,fPC2~Strata,lambda="ML")
fPC2mods[[3]]<-pgls(data=comp.data,fPC2~Migration,lambda="ML")
fPC2mods[[4]]<-pgls(data=comp.data,fPC2~Forest.Dependency,lambda="ML")
fPC2mods[[5]]<-pgls(data=comp.data,fPC2~nonforest+Strata,lambda="ML")
fPC2mods[[6]]<-pgls(data=comp.data,fPC2~nonforest+Migration,lambda="ML")
fPC2mods[[7]]<-pgls(data=comp.data,fPC2~Forest.Dependency+Strata,lambda="ML")
fPC2mods[[8]]<-pgls(data=comp.data,fPC2~Forest.Dependency+Migration,lambda="ML")
fPC2mods[[9]]<-pgls(data=comp.data,fPC2~Migration+Strata,lambda="ML")
fPC2mods[[10]]<-pgls(data=comp.data,fPC2~Migration+Strata+nonforest,lambda="ML")
fPC2mods[[11]]<-pgls(data=comp.data,fPC2~Migration+Strata+Forest.Dependency,lambda="ML")
fPC2mods[[12]]<-pgls(data=comp.data,fPC2~Habitat,lambda="ML")
fPC2mods[[13]]<-pgls(data=comp.data,fPC2~Habitat+Strata,lambda="ML")
fPC2mods[[14]]<-pgls(data=comp.data,fPC2~Habitat+Migration,lambda="ML")
fPC2mods[[15]]<-pgls(data=comp.data,fPC2~Migration+Strata+Habitat,lambda="ML")

# Collect summary statistics
fPC2modsAIC<-matrix(nrow=15,ncol=5)
for (i in 1:15){
  fPC2modsAIC[i,2]<-fPC2mods[[i]]$aicc
  fPC2modsAIC[i,3]<-summary(fPC2mods[[i]])$adj.r.squared
  fPC2modsAIC[i,4]<-pf(summary(fPC2mods[[i]])$fstatistic[1],summary(fPC2mods[[i]])$fstatistic[2],summary(fPC2mods[[i]])$fstatistic[3],lower.tail=F)
  fPC2modsAIC[i,5]<-fPC2mods[[i]]$param.CI$lambda$opt
  fPC2modsAIC[i,1]<-paste("Model",i,sep="_")
}

# transform to character
fPC2modsAIC<-data.frame(fPC2modsAIC)
fPC2modsAIC[,2]<-as.numeric(as.character(fPC2modsAIC[,2]))
fPC2modsAIC[,3]<-as.numeric(as.character(fPC2modsAIC[,3]))
fPC2modsAIC[,5]<-as.numeric(as.character(fPC2modsAIC[,5]))

# rename columns 
names(fPC2modsAIC)<-c("model","AICc","adj_r_sq","p_value","lambda")

# add plumage trait and habitat traits
fPC2modsAIC["Plumage_Trait"] <- c("fPC2")  ### change this out 
fPC2modsAIC["Habitat_Trait"] <- c(TRAITS)

fPC2modsAIC <- transform(fPC2modsAIC, `p_value` = as.numeric(`p_value`)) # convert p-value to numeric
fPC2modsAIC <- fPC2modsAIC %>% mutate(across(where(is.numeric), round, 3)) # round values to three decimal places

col_order <- c("Plumage_Trait", "Habitat_Trait", "model", "AICc", "adj_r_sq", "p_value", "lambda")
fPC2 <- fPC2modsAIC[, col_order]

fPC2<-fPC2[order(fPC2$AICc),]

write.csv(fPC2,"C:/Research/Masters_thesis/MT_tree_analysis/Results/PGLS/Chapter2/phylo_pca/fPC2_models.csv")
```

```{r brilliance pgls}
###########################################################################
## Brillance #########
# fAvgBrill
##############################
# Just categorical variables 
fAvgBrillmods<-list()
fAvgBrillmods[[1]]<-pgls(data=comp.data,fAvgBrill~nonforest,lambda="ML")
fAvgBrillmods[[2]]<-pgls(data=comp.data,fAvgBrill~Strata,lambda="ML")
fAvgBrillmods[[3]]<-pgls(data=comp.data,fAvgBrill~Migration,lambda="ML")
fAvgBrillmods[[4]]<-pgls(data=comp.data,fAvgBrill~Forest.Dependency,lambda="ML")
fAvgBrillmods[[5]]<-pgls(data=comp.data,fAvgBrill~nonforest+Strata,lambda="ML")
fAvgBrillmods[[6]]<-pgls(data=comp.data,fAvgBrill~nonforest+Migration,lambda="ML")
fAvgBrillmods[[7]]<-pgls(data=comp.data,fAvgBrill~Forest.Dependency+Strata,lambda="ML")
fAvgBrillmods[[8]]<-pgls(data=comp.data,fAvgBrill~Forest.Dependency+Migration,lambda="ML")
fAvgBrillmods[[9]]<-pgls(data=comp.data,fAvgBrill~Migration+Strata,lambda="ML")
fAvgBrillmods[[10]]<-pgls(data=comp.data,fAvgBrill~Migration+Strata+nonforest,lambda="ML")
fAvgBrillmods[[11]]<-pgls(data=comp.data,fAvgBrill~Migration+Strata+Forest.Dependency,lambda="ML")
fAvgBrillmods[[12]]<-pgls(data=comp.data,fAvgBrill~Habitat,lambda="ML")
fAvgBrillmods[[13]]<-pgls(data=comp.data,fAvgBrill~Habitat+Strata,lambda="ML")
fAvgBrillmods[[14]]<-pgls(data=comp.data,fAvgBrill~Habitat+Migration,lambda="ML")
fAvgBrillmods[[15]]<-pgls(data=comp.data,fAvgBrill~Migration+Strata+Habitat,lambda="ML")

# Collect summary statistics
fAvgBrillmodsAIC<-matrix(nrow=15,ncol=5)
for (i in 1:15){
  fAvgBrillmodsAIC[i,2]<-fAvgBrillmods[[i]]$aicc
  fAvgBrillmodsAIC[i,3]<-summary(fAvgBrillmods[[i]])$adj.r.squared
  fAvgBrillmodsAIC[i,4]<-pf(summary(fAvgBrillmods[[i]])$fstatistic[1],summary(fAvgBrillmods[[i]])$fstatistic[2],summary(fAvgBrillmods[[i]])$fstatistic[3],lower.tail=F)
  fAvgBrillmodsAIC[i,5]<-fAvgBrillmods[[i]]$param.CI$lambda$opt
  fAvgBrillmodsAIC[i,1]<-paste("Model",i,sep="_")
}

# transform to character
fAvgBrillmodsAIC<-data.frame(fAvgBrillmodsAIC)
fAvgBrillmodsAIC[,2]<-as.numeric(as.character(fAvgBrillmodsAIC[,2]))
fAvgBrillmodsAIC[,3]<-as.numeric(as.character(fAvgBrillmodsAIC[,3]))
fAvgBrillmodsAIC[,5]<-as.numeric(as.character(fAvgBrillmodsAIC[,5]))

# rename columns 
names(fAvgBrillmodsAIC)<-c("model","AICc","adj_r_sq","p_value","lambda")

# add plumage trait and habitat traits
fAvgBrillmodsAIC["Plumage_Trait"] <- c("fAvgBrill")  ### change this out 
fAvgBrillmodsAIC["Habitat_Trait"] <- c(TRAITS)

fAvgBrillmodsAIC <- transform(fAvgBrillmodsAIC, `p_value` = as.numeric(`p_value`)) # convert p-value to numeric
fAvgBrillmodsAIC <- fAvgBrillmodsAIC %>% mutate(across(where(is.numeric), round, 3)) # round values to three decimal places

col_order <- c("Plumage_Trait", "Habitat_Trait", "model", "AICc", "adj_r_sq", "p_value", "lambda")
fAvgBrill <- fAvgBrillmodsAIC[, col_order]

fAvgBrill<-fAvgBrill[order(fAvgBrill$AICc),]

write.csv(fAvgBrill,"C:/Research/Masters_thesis/MT_tree_analysis/Results/PGLS/Chapter2/phylo_pca/fAvgBrill_models.csv")
#############

# mAvgBrill
##############################
# Just categorical variables 
mAvgBrillmods<-list()
mAvgBrillmods[[1]]<-pgls(data=comp.data,mAvgBrill~nonforest,lambda="ML")
mAvgBrillmods[[2]]<-pgls(data=comp.data,mAvgBrill~Strata,lambda="ML")
mAvgBrillmods[[3]]<-pgls(data=comp.data,mAvgBrill~Migration,lambda="ML")
mAvgBrillmods[[4]]<-pgls(data=comp.data,mAvgBrill~Forest.Dependency,lambda="ML")
mAvgBrillmods[[5]]<-pgls(data=comp.data,mAvgBrill~nonforest+Strata,lambda="ML")
mAvgBrillmods[[6]]<-pgls(data=comp.data,mAvgBrill~nonforest+Migration,lambda="ML")
mAvgBrillmods[[7]]<-pgls(data=comp.data,mAvgBrill~Forest.Dependency+Strata,lambda="ML")
mAvgBrillmods[[8]]<-pgls(data=comp.data,mAvgBrill~Forest.Dependency+Migration,lambda="ML")
mAvgBrillmods[[9]]<-pgls(data=comp.data,mAvgBrill~Migration+Strata,lambda="ML")
mAvgBrillmods[[10]]<-pgls(data=comp.data,mAvgBrill~Migration+Strata+nonforest,lambda="ML")
mAvgBrillmods[[11]]<-pgls(data=comp.data,mAvgBrill~Migration+Strata+Forest.Dependency,lambda="ML")
mAvgBrillmods[[12]]<-pgls(data=comp.data,mAvgBrill~Habitat,lambda="ML", bounds = list(lambda=c(0.001,1), kappa=c(1e-6,3), delta=c(1e-6,3)))
mAvgBrillmods[[13]]<-pgls(data=comp.data,mAvgBrill~Habitat+Strata,lambda="ML")
mAvgBrillmods[[14]]<-pgls(data=comp.data,mAvgBrill~Habitat+Migration,lambda="ML")
mAvgBrillmods[[15]]<-pgls(data=comp.data,mAvgBrill~Migration+Strata+Habitat,lambda="ML")

# Collect summary statistics
mAvgBrillmodsAIC<-matrix(nrow=15,ncol=5)
for (i in 1:15){
  mAvgBrillmodsAIC[i,2]<-mAvgBrillmods[[i]]$aicc
  mAvgBrillmodsAIC[i,3]<-summary(mAvgBrillmods[[i]])$adj.r.squared
  mAvgBrillmodsAIC[i,4]<-pf(summary(mAvgBrillmods[[i]])$fstatistic[1],summary(mAvgBrillmods[[i]])$fstatistic[2],summary(mAvgBrillmods[[i]])$fstatistic[3],lower.tail=F)
  mAvgBrillmodsAIC[i,5]<-mAvgBrillmods[[i]]$param.CI$lambda$opt
  mAvgBrillmodsAIC[i,1]<-paste("Model",i,sep="_")
}

# transform to character
mAvgBrillmodsAIC<-data.frame(mAvgBrillmodsAIC)
mAvgBrillmodsAIC[,2]<-as.numeric(as.character(mAvgBrillmodsAIC[,2]))
mAvgBrillmodsAIC[,3]<-as.numeric(as.character(mAvgBrillmodsAIC[,3]))
mAvgBrillmodsAIC[,5]<-as.numeric(as.character(mAvgBrillmodsAIC[,5]))

# rename columns 
names(mAvgBrillmodsAIC)<-c("model","AICc","adj_r_sq","p_value","lambda")

# add plumage trait and habitat traits
mAvgBrillmodsAIC["Plumage_Trait"] <- c("mAvgBrill")  ### change this out 
mAvgBrillmodsAIC["Habitat_Trait"] <- c(TRAITS)

mAvgBrillmodsAIC <- transform(mAvgBrillmodsAIC, `p_value` = as.numeric(`p_value`)) # convert p-value to numeric
mAvgBrillmodsAIC <- mAvgBrillmodsAIC %>% mutate(across(where(is.numeric), round, 3)) # round values to three decimal places

col_order <- c("Plumage_Trait", "Habitat_Trait", "model", "AICc", "adj_r_sq", "p_value", "lambda")
mAvgBrill <- mAvgBrillmodsAIC[, col_order]

mAvgBrill<-mAvgBrill[order(mAvgBrill$AICc),]

write.csv(mAvgBrill,"C:/Research/Masters_thesis/MT_tree_analysis/Results/PGLS/Chapter2/phylo_pca/mAvgBrill_models.csv")
```
############### Taking the best models ###################

```{r best_model}

best_mpc1 <- mPC1[which.min(mPC1$AICc),]
best_mpc2 <- mPC2[which.min(mPC2$AICc),]
best_fpc1 <- fPC1[which.min(fPC1$AICc),]
best_fpc2 <- fPC2[which.min(fPC2$AICc),]
best_mbri <- mAvgBrill[which.min(mAvgBrill$AICc),]
best_fbri <- fAvgBrill[which.min(fAvgBrill$AICc),]

best_all_models <- rbind(best_mpc1,best_mpc2,best_fpc1,best_fpc2,best_mbri,best_fbri)
write_csv(best_all_models, file = "C:/Research/Masters_thesis/MT_tree_analysis/Results/PGLS/Chapter2/phylo_pca/best_models.csv")
```

c <- pgls(data=comp.data,mPC1~nonforest,lambda="ML")
summary(c)
c$aicc
c <- pgls(data=comp.data,mPC2~Habitat,lambda="ML")
summary(c)
c$aicc


```{r combine all_modles together}
all_Merged <- do.call("rbind", list(mPC1, mPC2, fPC1, fPC2, mAvgBrill, fAvgBrill))
write_csv(all_Merged, file = "C:/Research/Masters_thesis/MT_tree_analysis/Results/PGLS/Chapter2/phylo_pca/all_single_models.csv")

```

